### 2021-09-23 documentos/math.tex
Choi_N

\afnote{Notación para matrices densidad vectorizadas? $\varrho$ ?}
\janote{Qué tal $\vec\rho\quad$?}

We now turn to find the matrix form of the map $\mathcal{E}$ in order to construct the Choi matrix and the conditions the set of parameters $\{\tau_{\alpha}\}$ need to satisfy in order to $\mathcal{E}$ be considered a quantum channel. By writing the density matrix in vector form $\ket{\rho}=\sum_{\alpha=0}^{3} \alpha_{\alpha} \ket{\sigma_{\alpha}}$, it is easy to show that the matrix form of the map $\mathcal{E}$ may be written as
% 
\begin{equation}
\hat{\mathcal{E}} = \frac{1}{2}\sum_{\alpha=0}^{3} \tau_{\alpha} \dyad{\sigma_{\alpha}},
% \label{Map_Pauli}
\end{equation}
%
where the kets $\ket{\sigma_{\alpha}}$ correspond to the vectorized elements of
the Pauli basis and satisfy the usual relation $
\braket{{\sigma_{\alpha}} }{{\sigma_{\alpha'} }} = \tr \left(\sigma_{\alpha}^{\dagger} \sigma_{\alpha'}\right) = 2
~\delta_{\alpha\alpha'}$. After some steps it is possible to show that the
associated Choi matrix reads
% 
\begin{equation}
 \mathcal{D} = \frac{1}{2}\sum_{\alpha=0}^{3} \tau_{\alpha} \sigma_{\alpha} \otimes \sigma_{\alpha}^*.
%  \label{Choi_1}
\end{equation}
% 
Note that $\mathcal{D}$ has the same form as the matrix representation of a random unitary channel whose Kraus operators are given by $K_{\alpha}=\sqrt{\frac{\tau_{\alpha}}{2}}\sigma_{\alpha}$. For the latter case, it has been shown that corresponding eigenvalues read \cite{CHRUSCINSKI20131425}
% 
\begin{equation}
 \lambda_{\beta} = \frac{1}{2}\sum_{\alpha=0}^3H_{\beta\alpha}^{(4)}\tau_{\alpha},
\end{equation}
%
or
% 
\begin{equation}
 \vec{\lambda} = \frac{1}{2}~ H^{(2)}\otimes H^{(2)}\vec{\tau},
\end{equation}
% 
where $H^{(m)}$ is the Hadamard matrix of order \textit{m}, and for $m=2$ reduces to the usual Hadamard matrix.
% 
\begin{equation}
 H^{(2)}=\begin{pmatrix}
        1 & 1\\
        1 & -1\\
    \end{pmatrix}.
 \label{Hadd_Mat}
\end{equation}

% For the sake of simplicity in the derivation of the relations, let us employ a set of binary indices
% 
% \begin{equation}
% \rho=\sum_{mn=0}^{1} \alpha_{mn} \sigma_{mn},
% % \label{rho}
% \end{equation}

\subsubsection{$N$ qubits}
% 
For the case of $N$ qubits, the density matrix may be written as
% 
\begin{equation}
\rho=\sum_{\vec{\alpha}} r_{\vec{\alpha}} \sigma_{\vec{\alpha}}.
% \label{rho2}
\end{equation}
% 
Analogously to the single qubit case, we can define a $N$-qubits Pauli map
% 
\begin{equation}
\rho \to \rho'=\mathcal{E}[\rho ]=\sum_{\vec{\alpha}} \tau_{\vec{\alpha}} r_{\vec{\alpha}} \sigma_{\vec{\alpha}},
% \label{Pauli_Map}
\end{equation}
% 
with $-1\leq \tau_{\vec{\alpha}} \leq 1$. 
% 
In this case the density matrix in vector form holds 
% 
\begin{equation}
 \ket{\rho}=\sum_{\vec{\alpha}} r_{\vec{\alpha}} \ket{\sigma_{\vec{\alpha} }},
\end{equation}
% 
where $\ket{\sigma_{\vec{\alpha} }}=\ket{\sigma_{\alpha_1}}\otimes\dots \otimes\ket{\sigma_{\alpha_N}}$, and satisfy the usual orthogonality relation $\braket{{\sigma_{\vec{\alpha} } }}{{\sigma_{\vec{\alpha}'} }} = \tr \left(\sigma_{\vec{\alpha}}^{\dagger} \sigma_{\vec{\alpha}'}\right) = 2^N ~\delta_{\vec{\alpha}\vec{\alpha}'}$.
(Proofs in an appendix using binary indices XD)

% By writing the elements of the density matrix in vector form $\varrho=\left(\varrho_0,\varrho_1,\varrho_2,\varrho_3\right)^{\intercal}$, the action of a Pauli map on $\varrho$ for a single qubit may be written as $\varrho \to \varrho'=\hat \mcE \varrho$, where $\hat \mcE$ is the matrix representation of the map. Moreover it can be shown that in the Pauli basis $(\sigma_0,\sigma_1,i\sigma_2,\sigma_3)$, $\hat \mcE$ reduces simply to $\mathrm{diag}\left(\tau_0,\tau_1,\tau_2,\tau_3\right)$. 
% 
% Let reorganize the elements of the Pauli basis as $(\sigma_0,\sigma_3,\sigma_1,i\sigma_2)$, in this way it is possible to write the matrix representation of the map as $\hat \mcE = \tau_{03} \oplus \tau_{12}$, where $\tau_{jk}=\mathrm{diag}\left(\tau_j,\tau_k\right)$ and the symbol $``\oplus "$ represents a direct sum of matrices. The advantage of this ordering is manifested when the matrix $\hat \mcE$ is transformed to the computational basis $(u_{00},u_{11},u_{01},u_{10})$, $\hat \mcE_c$:
% % 
% \begin{align}
%  \hat \mcE_c =& \left(H\oplus H\right)\hat \mcE \left(H\oplus H\right)^{\intercal}\\
%  =& \left(H\oplus H\right) \left(\tau_{03} \oplus \tau_{12}\right)\left(H\oplus H\right)^{\intercal}\\
%  =& H\tau_{03}H \oplus H\tau_{12}H\\
%  =&\begin{pmatrix}
%         \tau_0+\tau_3 & \tau_0-\tau_3\\
%         \tau_0-\tau_3 & \tau_0+\tau_3\\
%         &&\tau_0+\tau_3 & \tau_0-\tau_3\\
%         &&\tau_0-\tau_3 & \tau_0+\tau_3\\        
%     \end{pmatrix},
%  \label{Eq_E_P_C}
% \end{align}
% % 
% where $H$ is the Haddamard matrix
% % 
% \begin{equation}
%  H=\begin{pmatrix}
%         1 & 1\\
%         1 & -1\\
%     \end{pmatrix}.
%  \label{Hadd_Mat}
% \end{equation}
% % 
% After applying the reshuffling operation on $\hat \mcE_c$, we get the associated Choi matrix (or sometimes called dynamical matrix) of the map $D$, it reads
% % 
% \begin{equation}
%  D=\begin{pmatrix}
%         \tau_0+\tau_3 & \tau_1+\tau_2\\
%         \tau_1+\tau_2 & \tau_0+\tau_3\\
%         &&\tau_0-\tau_3 & \tau_1-\tau_2\\
%         &&\tau_1-\tau_2 & \tau_0-\tau_3\\        
%     \end{pmatrix}.
%  \label{Eq_D_P}
% \end{equation}
% % 
% Back to the Pauli basis, the Choi matrix holds
% \begin{align}
%  D_p =& \left(\frac{1}{2}H\oplus H\right)\hat D \left(\frac{1}{2}H\oplus H\right)^{\intercal}\\
%  =&\frac{1}{4}\left(H\oplus H\right)\hat D \left(H\oplus H\right)^{\intercal}\\
% =&\begin{pmatrix}
%         \lambda_0\\
%         &\lambda_1\\
%         &&\lambda_2\\
%         &&&\lambda_3\\
%     \end{pmatrix}.
% %   =&\begin{pmatrix}
% %         \tau_0+\tau_3+\tau_1+\tau_2\\
% %         &\tau_0-\tau_3+\tau_1-\tau_2\\
% %         &&\tau_0+\tau_3-\tau_1-\tau_2\\
% %         &&&\tau_0-\tau_3-\tau_1+\tau_2\\        
% %     \end{pmatrix}.
%  \label{Eq_D_C_P}
% \end{align}
% % 
% In other words, the Pauli basis diagonalizes the Choi matrix associated to any Pauli map, with eigenvalues related to the map coefficients $\tau$ as
% % 
% \begin{equation}
% \begin{pmatrix}
%         \lambda_0\\
%         \lambda_1\\
%         \lambda_2\\
%         \lambda_3\\   
%     \end{pmatrix}
%  =\frac{1}{2}\begin{pmatrix}
%         1&1&1&1\\
%         1&-1&1&-1\\
%         1&1&-1&-1\\
%         1&-1&-1&1\\
%     \end{pmatrix}
%     \begin{pmatrix}
%         \tau_0\\
%         \tau_3\\
%         \tau_1\\
%         \tau_2\\   
%     \end{pmatrix},
%  \label{Eigen1}
% \end{equation}
% % 
% or in a compact form
% % 
% \begin{equation}
%  \lambda = \frac{1}{2}H\otimes H \tau.
% \end{equation}
% % 
% Now we proceed to generalize this result for the case $N$ qubits case.

% \subsubsection{$N$ qubits}

>>>>

\afnote{Notación para matrices densidad vectorizadas? $\varrho$ ?}
\janote{Qué tal $\vec\rho\quad$?}
\afnote{Por ahora uso este $\ket{\rho}$}

We now turn to find the matrix form of the map $\mathcal{E}$ in order to construct the Choi matrix and the conditions the set of parameters $\{\tau_{\alpha}\}$ need to satisfy in order to $\mathcal{E}$ be considered a quantum channel. By writing the density matrix in vector form $\ket{\rho}=2^{-1}\sum_{\alpha=0}^{3} r_{\alpha} \ket{\sigma_{\alpha}}$, the matrix form of the map $\mathcal{E}$ may be written as
% 
\begin{equation}
\hat{\mathcal{E}} = \frac{1}{2}\sum_{\alpha=0}^{3} \tau_{\alpha} \dyad{\sigma_{\alpha}},
% \label{Map_Pauli}
\end{equation}
%
where the kets $\ket{\sigma_{\alpha}}$ correspond to the vectorized elements of the Pauli basis and satisfy the usual relation $ \braket{{\sigma_{\alpha}}}{{\sigma_{\alpha'}}} = \tr \left(\sigma_{\alpha}^{\dagger} \sigma_{\alpha'}\right) = 2 ~\delta_{\alpha\alpha'}$. After some steps it is possible to show that the associated Choi matrix reads
% 
\begin{equation}
 \mathcal{D} = \frac{1}{2}\sum_{\alpha=0}^{3} \tau_{\alpha} \sigma_{\alpha} \otimes \sigma_{\alpha}^*.
%  \label{Choi_1}
\end{equation}
% 
Moreover, observe that $\mathcal{D}$ has the same form as the matrix representation of a random unitary channel, with Kraus operators given by $K_{\alpha}=\sqrt{\tau_{\alpha}/2}\sigma_{\alpha}$ \cite{CHRUSCINSKI20131425}. In this way the Choi matrix $\mathcal{D}$ is diagonal in the Pauli basis (see Appendix for details),
%
\begin{equation}
 \mathcal{D} \ket{\sigma_{\alpha}}= \lambda_{\alpha} \ket{\sigma_{\alpha}},
\end{equation}
% 
with eigenvalues given by 
% 
\begin{equation}
 \lambda_{\alpha} = \sum_{\beta=0}^3\sa_{\alpha\beta}\tau_{\beta},
\end{equation}
%
or more compactly $\vec{\lambda} = \sa~\vec{\tau}$, where $\sa=2^{-1}H\otimes H$ and $H$ is the Hadamard matrix,
%
\begin{equation}
 H=\begin{pmatrix}
        1 & 1\\
        1 & -1\\
    \end{pmatrix}.
 \label{Hadd_Mat}
\end{equation}
% 
\afnote{Podemos definir la matriz de Hadamard con un factor $2^{-1/2}$, o sea $H'=H/\sqrt{2}$ para evitar cargar factores extra hasta el final, de esta manera tenemos simplemente $\san=H^{\otimes 2N}$. Qué opinan?}

\subsubsection{$N$ qubits}
% 
Analogously to the single qubit case, one can define the $N$-qubits Pauli map $\mathcal{E}_N$
% 
\begin{equation}
\rho \to \rho'=\mathcal{E}_N[\rho ]=\frac{1}{2^N}\sum_{\vec{\alpha}} \tau_{\vec{\alpha}} r_{\vec{\alpha}} \sigma_{\vec{\alpha}},
% \label{Pauli_Map}
\end{equation}
% 
with $-1\leq \tau_{\vec{\alpha}} \leq 1$. 
% 
In this case the density matrix in vector form holds 
% 
\begin{equation}
 \ket{\rho}=\frac{1}{2^N}\sum_{\vec{\alpha}} r_{\vec{\alpha}} \ket{\sigma_{\vec{\alpha}}},
\end{equation}
% 
where $\ket{\sigma_{\vec{\alpha}}}=\ket{\sigma_{\alpha_1}}\otimes\dots \otimes\ket{\sigma_{\alpha_N}}$, and satisfy the usual orthogonality relation $\braket{{\sigma_{\vec{\alpha}}}}{{\sigma_{\vec{\alpha}'}}} = \tr \left(\sigma_{\vec{\alpha}}^{\dagger} \sigma_{\vec{\alpha}'}\right) = 2^N ~\delta_{\vec{\alpha}\vec{\alpha}'}$. The matrix representation of this map is given by
% 
\begin{equation}
\hat{\mathcal{E}}_N = \frac{1}{2^N}\sum_{\vec{\alpha}} \tau_{\vec{\alpha}} \dyad{\sigma_{\vec{\alpha}}}.
% \label{Map_Pauli}
\end{equation}
%
As in the previous case, the Choi matrix $\mathcal{D}_N$ may be written in terms of tensor products of Pauli matrices
% 
\begin{equation}
 \mathcal{D}_N = \frac{1}{2^N}\sum_{\vec{\alpha}} \tau_{\vec{\alpha}}  \bigotimes_{j=1}^N \sigma_{\alpha_j} \otimes \sigma_{\alpha_j}^*.
%  \label{Choi_1}
\end{equation}
% 
After several steps, it is possible to prove that analogously to the single qubit case, $\mathcal{D}_N$ is diagonal in the $N$-qubits Pauli basis
%
\begin{equation}
 \mathcal{D}_N \ket{\sigma_{\vec{\alpha}}}= \lambda_{\vec{\alpha}} \ket{\sigma_{\vec{\alpha}}},
\end{equation}
% 
with eigenvalues $\lambda_{\vec{\alpha}}$ given by 
% 
\begin{equation}
 \lambda_{\vec{\alpha}} = \sum_{\vec{\beta}} \san_{\vec{\alpha}\vec{\beta}}\tau_{\vec{\beta}},
\end{equation}
%
where 
% 
\begin{equation}
 \san=\sa^{\otimes N}=\frac{1}{2^N}\left(H\otimes H\right)^{\otimes N}=\left(\frac{H}{\sqrt{2}}\right)^{\otimes 2N}.
\end{equation}
%

(Proofs in an appendix using binary indices XD)

\subsubsection{Appendix - Diagonalization of Choi Matrix $\mathcal{D}_N$}
